---
title: "p8105_hw2_zx2373"
author: Ziyan Xu
date: 09/28/2022
output: github_document
---

```{r message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

### Read and clean the NYC Transit dataset. 

```{r message = FALSE}
nyctransit_df = 
  read_csv(
    "NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  rename(
    ada_compliance = ada
    ) %>% 
  dplyr::select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada_compliance
    ) %>% 
  mutate(
    entry = case_when(entry == "YES" ~ "TRUE", entry == "NO" ~ "FALSE"), 
    entry = as.logical(entry)
    )
```

- This dataset contains information of the subway transformation in NYC, related to entrance and exit for each station. The dataset contains information on line, station name and location (longitude, latitude), routes served, entry, vending, entrance type, and ADA compliance. 

- Firstly, I load the dataset into nyctransit_df, clean the variable names and rename the 'ada' and select variables of interest.

- Then I convert the entry variable from character (YES vs NO) to a logical variable(TRUE vs FALSE). It is not a completely tidy format, since there are several columns of "routes_served" that can be addressed into more clear format (in the following text).

- To sum up, there are a total of `r nrow(nyctransit_df)` rows, `r ncol(nyctransit_df)` columns in the final dataset.

Other related questions:

- There are `r nyctransit_df %>% distinct(station_name, line) %>% nrow` distinct stations, identified both by name and by line.

- There are `r nyctransit_df %>% filter(ada_compliance == "TRUE") %>% nrow` stations are ADA compliant.

- The proportion of station entrances / exits without vending allow entrance is `r  nyctransit_df %>% filter(vending == "NO" & entry == "TRUE") %>% nrow / nyctransit_df %>% filter(vending == "NO") %>% nrow`

### Reformat data so that route number and route name are distinct variables: 

```{r}
nyctransit_df_tidy = 
  nyctransit_df %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "routes_name",
    names_prefix = "route",
    values_to = "routes_number"
    ) %>% 
  drop_na(routes_number) 
```

- There are  `r nyctransit_df_tidy %>% distinct(station_name, line, .keep_all = TRUE) %>% filter(routes_number == "A") %>% nrow`  distinct stations serve the A train.

- Of the stations that serve the A train, there are `r nyctransit_df_tidy %>% distinct(station_name, line, .keep_all = TRUE) %>% filter(routes_number == "A" & ada_compliance == "TRUE") %>% nrow` ADA compliant.


## Problem 2

### Read the Mr. Trash Wheel dataset.

```{r}
mr_df = 
  read_xlsx(
    "Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  filter(dumpster != "Grand Total") %>%
  mutate(sports_balls = as.integer(round(sports_balls,0)),
         dumpster = as.integer(dumpster)) %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards)
view(mr_df)
```

### Read the Professor Trash Wheel dataset.

```{r}
prof_df = 
  read_xlsx(
    "Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  dplyr::select(-x14)
view(prof_df)
```

### Combine two datasets.

```{r}
mr_merge_df = 
  mutate(mr_df, Trash_wheel = c('Mr'), dumpster = as.character(dumpster))

prof_merge_df = 
  mutate(prof_df,
         Trash_wheel = c('Pro'),
         dumpster = as.character(dumpster),
         year = as.character(year))

merge_df = bind_rows(mr_merge_df, prof_merge_df) %>% 
  relocate(Trash_wheel, .before = "weight")

view(merge_df)
show(merge_df)
```

### Total weight of trash collected by Professor Trash Wheel

```{r}
total_weight = sum(prof_df$weight) %>% 
show()
```

### Total number of sports balls collected by Mr. Trash Wheel in 2020

```{r}
total_ball = mr_df %>% 
  filter(year == "2020") %>% 
  pull(sports_balls) %>% 
  sum() %>% 
show()
```

- There are `r nrow(merge_df)` observations in the merged dataset, including `r ncol(merge_df)` variables, namely `r names(merge_df)`.

- The total weight of trash collected by Professor Trash Wheel is `r total_weight` tons, and the total number of sports balls collected by Mr. Trash Wheel in 2020 was `r total_ball`.


## Problem 3

### Clean the data in pols-month.csv

```{r}
Pols_month_df = read_csv(
  "./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>% 
  drop_na() %>% 
  separate(mon, c("year","month","day")) %>%
  mutate(
    year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%
  mutate(month = month.abb[month]) %>% 
  mutate(president = ifelse(prez_gop == "0", "dem","gop")) %>% 
  relocate(president, .before = "gov_gop") %>% 
  dplyr::select(-prez_dem, -prez_gop, -day)
```

```{r}
show(Pols_month_df)
```

### Clean the data in snp.csv

```{r}
snp_df = read_csv(
  "./fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  janitor::clean_names() %>% 
  mutate(
    date = lubridate::parse_date_time2(date,orders = "mdy", cutoff_2000 = 23)) %>% 
  separate(date, c("year","month","day")) %>%
  mutate(
    year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%
  mutate(month = month.abb[month]) %>% 
  relocate(year, .before = "month") %>% 
  dplyr::select(-day)
```

```{r}
show(snp_df)
```

### Clean the data in unemployment.csv

```{r}
unemploy_df = read_csv(
  "./fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,names_to = "month", values_to = "unemployment_rate") %>%
  janitor::clean_names() 
```

```{r}
show(unemploy_df)
```

### Join the datasets

```{r}
snp_pols = 
  left_join(Pols_month_df, snp_df, by = c("year", "month")) 

merge_df = 
  left_join(snp_pols, unemploy_df, by = c("year", "month")) %>% 
  view()
```

pols-month data :

The data set describes the political distribution in the United State seperately in each year from 1947 to 2015 with 822 rows and 9 columns after cleaning. Variable President means whether the president is Republican or Democratic.

snp data :

The snp data set represents the closing values of the S&P stock index on the associated date from 1969 to 2068 with 787 rows and 2 columns. Year means the yaer of observation. Month means the month of observation. close means the closing calue.

unemployment data :

The unemployment data contains the unemployment rate in each month from 1948 to 2014 with 804 rows and 3 columns after changed to cleaned version. Year means the yaer of observation. Month means the month of observation. Percentage means unemployment rate.

Combined dataset:

This dataset combines the former three data on the scale of year and month. It shows data in economics(stock and unemployment rate) and politics(political distribution) in one sheet with 822 rows and 11 columns from 1947 to 2015.



