---
title: "p8105_hw2_zx2373"
author: Ziyan Xu
date: 09/28/2022
output: 
  github_document:
    toc: true
---

```{r message = FALSE}
library(tidyverse)
library(readxl)
```

#Problem 1

###Read and clean the NYC Transit dataset. 

```{r message = FALSE}
nyctransit_df = 
  read_csv(
    "NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  rename(
    ada_compliance = ada
    ) %>% 
  select(
    line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada_compliance
    ) %>% 
  mutate(
    entry = case_when(entry == "YES" ~ "TRUE", entry == "NO" ~ "FALSE"), 
    entry = as.logical(entry)
    )
```

- This dataset contains information of the subway transformation in NYC, related to entrance and exit for each station. The dataset contains information on line, station name and location (longitude, latitude), routes served, entry, vending, entrance type, and ADA compliance. 

- Firstly, I load the dataset into nyctransit_df, clean the variable names and rename the 'ada' and select variables of interest.

- Then I convert the entry variable from character (YES vs NO) to a logical variable(TRUE vs FALSE). It is not a completely tidy format, since there are several columns of "routes_served" that can be addressed into more clear format (in the following text).

- To sum up, there are a total of `r nrow(nyctransit_df)` rows, `r ncol(nyctransit_df)` columns in the final dataset.

Other related questions:

- There are `r nyctransit_df %>% distinct(station_name, line) %>% nrow` distinct stations, identified both by name and by line.

- There are `r nyctransit_df %>% filter(ada_compliance == "TRUE") %>% nrow` stations are ADA compliant.

- The proportion of station entrances / exits without vending allow entrance is `r  nyctransit_df %>% filter(vending == "NO" & entry == "TRUE") %>% nrow / nyctransit_df %>% filter(vending == "NO") %>% nrow`

###Reformat data so that route number and route name are distinct variables: 

```{r}
nyctransit_df_tidy = 
  nyctransit_df %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "routes_name",
    names_prefix = "route",
    values_to = "routes_number"
    ) %>% 
  drop_na(routes_number) 
```

- There are  `r nyctransit_df_tidy %>% distinct(station_name, line, .keep_all = TRUE) %>% filter(routes_number == "A") %>% nrow`  distinct stations serve the A train.

- Of the stations that serve the A train, there are `r nyctransit_df_tidy %>% distinct(station_name, line, .keep_all = TRUE) %>% filter(routes_number == "A" & ada_compliance == "TRUE") %>% nrow` ADA compliant.


#Problem 2

###Read the Mr. Trash Wheel dataset.

```{r}
mr_df = 
  read_xlsx(
    "Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

###Read the Professor Trash Wheel dataset.

```{r}
prof_df = 
  read_xlsx(
    "Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "Professor Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

###Combine two datasets.

```{r}

```


#Problem 3

###Clean the data in pols-month.csv

```{r}

```

